---
title: "PH251D Fall 2019 - Project 2 (v. 2019-12-01-A)"
author: "FirstName MI LastName"
date: "2019-MM-DD"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NULL)
```

Download this Rmarkdown template (`PH251D2019_LastName_Project2.Rmd`)
and edit.

Create a project folder called `project2` on your computer. You will
put all your Project 2 files in this folder.

Use this template and R Markdown to demonstrate the skills below. Be
sure to compile this document to .DOC, .HTML, or PDF file for
submission. Be sure to change `output` option in YAML.

You will need to install `bnlearn`, `tableone`, `survey`, and
`sandwich` packages.

## 1. Reading in CSV file

The CAGE is a four-item questionnaire for screening for alcohol use
disorder. A patient can answer "yes" to zero, one, two, three, or four
questions. Mayfield conducted a CAGE study in 1974 with 366 patients,
recorded the number of CAGE questions that were answered "Yes" (range
was 0 to 4). A psychiatric social worker independently conducted
in-depth interviews and each patient was classified by whether he or
she had alcohol use disorder or not (AUD: "Y", "N"). The AUD status
was considered the gold standard. The study results are available
here:
<https://raw.githubusercontent.com/taragonmd/data/master/cage.csv>

Read in the data set and display a two-way continency table using the
`xtabs` function. The column headings should be the `AUD` status.

```{r}
## R code here to display two-way table similar to above
```

## 2. Recode integer variable into a factor with two levels

Recode the `num_yes` variable into a new factor variable called
`CAGE`:

| Factor | Level | `num_yes` values |
|:------:+:-----:+:----------------:|
| `CAGE` | `pos` | >= 3             |
|        | `neg` | <= 2             |

Use the `xtabs` function to display 2 by 2 table of CAGE by AUD.

```{r}
## display 2 x 2 table
```

## 3. Calculate the sensitivity and specificity of the CAGE questionnaire

The sensitivity is P(CAGE = pos | AUD = Y), and the specificity is
P(CAGE = neg | AUD = N). 


```{r}
## calculate sensitivity and specificity
```


## 4. Posterior probabilities as a function of prior probabilities

You have learned that the posterior (post-test) probability is a
function of the prior (pre-test) probability, and sensitivity and
specificity of the diagnostic test (in this case, the CAGE
questionnaire).  For alcohol use disorder (AUD), the positive
predictive value (PPV) is the probability that a patient has AUD given
that they answered "yes" to 3 or 4 items on the CAGE questionnaire.

Using the sensivity and specificity you calculated above, plot a graph
displaying on how PPV changes as a function of the prior
probability. Some of the R code is provided, including Bayes' theorem
for calculating PPV. 

```{r, eval=F}
## plot of PPV vs prior probability (prior)
## use `sens` and `spec` from previous results
prior <- seq(from = 0, to = 1, by 0.05)
ppv <- (sens * prior) / (sens * prior + (1 - spec) *(1 - prior))
```


## 5. Bayesian network for probabilistic reasoning

Install the `bnlearn` package as described here: <http://www.bnlearn.com/>.
The causal model is AUD -> CAGE. Create a Bayesian network (BN) model
by fitting the causal model to data (I have done this for you).

```{r, fig.width=3, fig.height=2, fig.align="center"}
library(bnlearn)
curl <- 'https://raw.githubusercontent.com/taragonmd/data/master/cage2.csv'
aud <- read.csv(curl, header = TRUE)
summary(aud)
dag <- empty.graph(nodes=c("AUD", "CAGE")) # Build the Bayesian network
dag <- set.arc(dag, from = "AUD", to = "CAGE")
graphviz.plot(dag, layout = "circo") # Plot the BN
bn <- bn.fit(dag, data = aud)        # Fit the BN to the data
# cpquery
```

Assuming the prior probability of AUD is 5%, use the `cpquery`
function from `bnlearn` package to calculate the PPV; that is P(AUD =
"Y" | CAGE = "pos").

## 6. Deconfounding: IPW with stratified contingency table

In 1986, Charig published a retrospective observational study
comparing open surgery (Treatment A) to percutaneous nephrolithotomy
(Treatment B) for the treatment of kidney stones. Treatment B is
noninvasive. Success was defined as no stones at three months. The
data are presented in Table below.

Table: Comparison of Treatment A to B for kidney stones, by size

| Stone size   | Treatment A | Success (%) |   | Treatment B | Success (%) |
|--------------+------------:+------------:+---+------------:+------------:|
| Small stones | 81/87       | **93**      |   | 234/270     |          87 |
| Large stones | 192/263     | **73**      |   | 55/80       |          69 |
| Combined     | 273/350     | 78          |   | 289/350     |      **83** |


Urologists select open surgery (Treatment A) for more severe cases, and
larger stones are considered more severe. In other words, severity
causes treatment selection, and severity lowers success. Stone size
(severity) is a common cause or fork that may be confounding the
causal relationship between treatment and success (outcome).

Review Section 4.3 from <https://escholarship.org/uc/item/8000r5m5>

The study results are available here:
<https://github.com/taragonmd/data/blob/master/kidney.csv>

Read in the data set. Using the inverse probability weighting (IPW) method
(see Section 4.3.1.1) and report on the average causal effect (ACE) in
terms of the following:

(a) causal risk difference
(b) causal risk ratio
(c) causal odds ratio

```{r}
## put R code here
```


## 7. Deconfounding: IPW with marginal structural model (MSM)

Repeat the analysis above using IPW/marginal structural model (MSM)
that using the propensity scores and the generalized linear model
(`glm` function in R). Edit the R code from Section 4.3.1.2 and report
on the average causal effect (ACE) in terms of the following:

(a) causal risk difference
(b) causal risk ratio
(c) causal odds ratio

You will need to install the `tableone`, `survey`, and `sandwich`
package.


```{r}
## Put R code here
```


